Molecular dynamics simulation of synchronization of a driven particle

### 2021-05-04 manuscript/manuscript_ajp_version4b.tex
cleaned up, added ignore

\begin{abstract}
  Synchronization
  plays an important role in many physical processes.
  We discuss synchronization in a 
  molecular dynamics simulation
  of a single particle
  moving through
  a viscous liquid
  while being driven 
  across a washboard potential energy landscape.
  Our results show many dynamical patterns
  as the landscape and driving force are altered.
  For certain conditions,
  the particle velocity and location
  are synchronized or 
  phase-locked,
  forming closed orbits in phase space.
  Quasi-periodic motion is common, 
  for which the
  dynamical center of motion shifts the
  phase space orbit.
  %DM removed first sentence,
  %which was strong and clear, but isolated in the abstract
  %put those ideas in this final sentence, making it a mouthful [26 words]
  By 
  isolating
  synchronized motion
  in 
  simulations and table-top experiments,
  we 
  can study %DM replaced isolate with study
  complex natural behaviors
  important to many physical processes.
 %[suggest omitting this sentence. Code not in text. We include molecular dynamics code
 % to simulate and characterize the
  %particle dynamics.] 
\end{abstract}

@@

\section{Introduction} 
Synchronization is a universal phenomena
in which
individual oscillators change frequency due
to external stimuli.\cite{Pikovsky2003}
The
flickering patterns of
candle flames 
mediated by temperature fluctuations,\cite{Okamoto2016}
the vibrations of singing wine glasses interacting 
through sound waves,\cite{Arane2009}
and metronomes vibrating through a supporting platform\cite{Jia2015}
are examples of in-phase environmentally coupled oscillations. 
Biological systems benefit from cooperative
synchronization --
birds  coordinate wing flaps
to optimize energy use during flight,\cite{Portugal2014}
frogs alternate croaking patterns,\cite{Aihara2014}
humans clap in time with music,\cite{Tranchant2016}
and 
at a cellular level, 
neurons simultaneously fire in cardiac muscle\cite{MartinHall1999}
and brain tissue.\cite{Singer1999}
%DM ended paragraph with fewer examples, 
%Environmental conditions
%DM add add transition sentence
%
External forcing
can cause or regulate synchronization.
%where external forces
%add or remove energy from the system.
For example, 
an electrical pacemaker 
regulates a heart beat and
a pulsed light modifies the
flashing pattern of fireflies.

The model discussed in this paper
resembles an externally forced pendulum.
%damping
%transition
%%DM - focus more on single particle with environment
A single pendulum 
can swing in time with an %be driven or damped by
driving force
like  %a skateboarder in a half-pipe or
a pushed child on a swing,
and may be damped
by mechanical friction.
%of gravity to return
%to an equilibrium position,
%or  to decreasing amplitude
% or air resistance.
We focus on 
synchronized phase-locking or mode-locking,
which 
first appeared in the scientific literature
in 1665 with 
Huygens'  experiments on the
motions of %synchronized 
pendula in wall-mounted clocks.\cite{Bennett2002}
A locked-mode is an integer frequency ratio
in coupled oscillators. %Clock pendula swing in phase
%when coupled through shared vibrations.
In Huygens' clocks,
the mode-locked pendula were observed to swing at the same rate
in the same direction (1:1 mode) or
opposite directions ($-1$:1 mode).
Higher modes might be achieved
by altering the pendula lengths - 
a 2:1 mode occurs when a simple pendulum
swings with twice the period
of another one quarter the length.
%DM transition sentence
%Many subsequent 
Simulations and laboratory experiments
have revealed 
mode-locked synchronization
in a diverse range of 
oscillation phenomena.
%Single oscillators 
%can demonstrate a variety of synchronized behaviors.
%Colloid particles in optical traps are an excellent
%model system to explore %these 
%the dynamics of mode-locking. %synchronization.

Colloids have been
shown to 
provide an ideal system to 
control and study
phase-locked motions
%are useful for
%understanding synchronization
at a single particle level.\cite{Juniper2015,Juniper2017}
% such as synchronized mode-locking can be studied with
%An advantage of colloid studies
In experiments %DM moved the ``in experiments'' to second sentence
typical colloids are   
plastic spheres suspended in
de-ionized water or silica beads suspended in organic solvent.
Colloids can be trapped
with 
radiation pressure from 
a laser beam.\cite{Ashkin1997} 
%A colloid in such a confining region,
%an optical trap,
A colloid centered in an optical trap is 
uniformly bombarded by photons
and oscillates
about a equilibrium point.
Off-center colloids 
experience a net force
due to uneven photon collisions across
the particle surface.
%Depending on the 
%location of the particle in the trap,
%the radiation pressure either moves colloids toward the center 
%or ejects it from the trap.
Diffraction gratings can be used to create
more complex light environments, 
such as periodic patterns of minima
suitable for synchronization studies.\cite{Grier2003}  
A driven colloid
can demonstrate
synchronized
motion within one trap,
or synchronized hopping between multiple traps. 
While many systems oscillate,
trapped colloids are useful for 
providing controlled 
measurements of the
step-by-step dynamics of particle motion.
Since
colloids are large 
and move slowly, 
%colloids are considered model systems
particle positions 
can be measured in real time 
with an optical camera.\cite{Pertsinidis2001}
%Light is a
%tool for manipulating the colloidal environment,
%which can alter synchronization patterns.
%Synchronization patterns can be altered
%using light to manipulate
%the colloidal environment.
Simulations provide complementary
information,
where environmental and driving parameters may be precisely
and quickly modified to explore
the many oscillation patterns.
%the dynamics from external
In the following study,
we confine colloids  
to move along a periodic pattern
with minima and maxima defined by a sine wave.
%a washboard potential energy substrate.

A colloid driven across a periodic
substrate may 
synchronize its location
to the pattern of the external drive.
%moving back and forth in time with
%the beat, or
%moving between substrate minima.
When 
a constant or 
dc drive is applied,
the particle velocity is modulated by 
the potential energy landscape exerted on the particle.   
Below some threshold  
the dc force is not strong enough to push the particle
across a potential maximum so the average velocity is zero,
a phenomena referred to as pinning.\cite{Reichhardt2017}
Above the pinning threshold,
a dc driven particle
%a particle subject to a constant
%drive force
increases its speed at a rate proportional
to the external drive.  
When the applied force varies periodically,
an ac drive,
%cause
the particle can hop back and forth across
the landscape minima.
When a particle is driven across a substrate
by combined 
ac and dc forces,
it is possible for the particle to demonstrate 
mode-locking.
In these modes,  
%Due to the combination of external drive
%and substrate minima,
the particle hops at a controlled rate such that 
the average particle velocity is 
%of an  driven particle 
fixed for a range of dc drive forces
with a magnitude proportional to the
landscape period and driving frequency.\cite{Reichhardt2015}
%Many synchronized patterns occur,
%controlled by the substrate pattern and driving force.

In this paper we discuss numerical simulations 
of the mode-locked dynamics
of a confined particle driven over
a washboard shaped potential energy landscape.
The numerical model is presented in Sec.~\ref{sec:MD} and 
our results are summarized in
%including the synchronized motion of a single confined particle
%driven across a periodic landscape in 
Sec.~\ref{sec:results}. In Sec.~\ref{sec:conclusion}
we relate  our results 
to a broad variety of physical systems beyond 
optically trapped colloids.
%and superconducting vortices in the presence of periodic pinning arrays.
We include problems for interested readers 
in Sec.~\ref{sec:problems}.

\section{Simulation}
\label{sec:MD}
We use a classical model for 
studying the dynamics, 
using the net force on a particle to calculate
its trajectory.
The particle is confined in a one-dimensional %DM reduced to 1D
simulation of length $L=46.\bar{6}\,a_0$, %DM 1D
where $a_0$ is a dimensionless unit of length.
(In the simulations we set $a_0 = 1$.)
The particle 
has
position $\vec{r} = y \hat{y}$ %DM 1D
and velocity $\vec{v} = d\vec{r}/dt$.
The edges of the system are treated by
periodic boundary conditions,
such that a particle leaving the edge of the system is mapped
back to a position within the simulation boundaries 
by the transformation $y+L \rightarrow y$. %DM 1D
We show a schematic of the system in Fig.~\ref{fig:1_landscape}(a).
The units of the simulated variables are summarized in Table~\ref{tab:1}.

We confine the particles using a position dependent 
potential energy, called a landscape or substrate.
The landscape is modulated in the $y$-direction
by the periodic function 
 \begin{equation}
   U(y) = U_0 \cos{(2 \pi y / \lambda)},
     \label{eq:ysubstrate}
\end{equation}
where $\lambda=L/N_p$, with $N_p$ equal to the number of periods,
and $U_0$ is a parameter
 that sets the depth of the minima
 with simulation units of energy $E_0$. 
 We plot this function in 
 Fig.~\ref{fig:1_landscape}
 for $N_p = 3$.  In Fig.~\ref{fig:1_landscape}(a) we show 
 the $x$-$y$ plane with a contour plot of $U(y)$ 
 to illustrate
 the 2D potential energy landscape;
 the maxima are colored gray and the minima colored white.

The confining force on the particle %$i$ DM removed
 is given by
 \begin{equation}
 \vec{F}_{\ell}(\vec{r}) = - \vec \nabla U(\vec{r}),
 \label{eq:dudr}
 \end{equation}
where $\ell$ denotes the landscape.
 In Fig.~\ref{fig:1_landscape}(b) we plot  
 $U(y)$ to illustrate how the magnitude
 $|\vec{F}_{\ell}|$ is calculated from the particle position $y$.

@@

The equation of motion provides a direct calculation of the velocity
of an individual particle at position $\vec{r}$ 
and simulation time $t$. At each time step
we evaluate the net force on the particle as a function of its position
$\vec{r}(t)$
and then integrate
the equation of motion to move the particle
to its updated position.
Because the acceleration is zero,
the integration of the equation of motion
is performed using 
the simple first-order Euler method 
\begin{equation}
  \vec{r}(t+\Delta t) = \vec{r}(t) + \vec{v}(t) \Delta t
    \label{eq:euler}
\end{equation}
for a time step $\Delta t = 0.1\,\tau$.
In 
Problem~\ref{ex:euler}
we describe 
the numerical methods for 
solving differential equations such as Eq.~\eqref{eq:euler}.
%------------------------------------------------------------------
\begin{figure} % [h!]
\centering
\includegraphics[width=\columnwidth]{fig1_landscape.pdf}
\caption{
Schematic of the simulation of a single particle
  driven across a washboard potential energy landscape.
  (a) View of the $x$-$y$ plane. 
  The time-dependent applied driving force $\vec{F}_{\rm d}$
  is parallel to the $y$-axis.
  The landscape is shown with
  maxima of the potential energy in gray 
  and minima in white. 
  The particle is 
  subject to competing forces of the landscape and the applied driving force.
  (b) The potential energy function $U(y)$. 
  The particle in (a) is shown at the same $y$-position in (b).
  The slope of $U(y)$ is the 
  magnitude of the force $\vec{F}_{\ell}$. 
  }
\label{fig:1_landscape}
\end{figure}

@@

\label{sec:results}
Here we drive 
a single particle 
across the landscape. 
The numerical implementation of the force due to the landscape 
is calculated using Eqs.~\eqref{eq:ysubstrate} and \eqref{eq:dudr}  
\begin{equation}
  \label{eq:force}
  {F_{\ell}}_y(y) = A_{p} \sin{(2 \pi y / \lambda)},
\end{equation}
where the force is scaled by the parameter $A_{p} = 2\pi U_0/\lambda$.
In this section we fix the landscape parameters
to $A_{p} = 0.1\,F_0$ 
with $N_p=20$ minima, 
corresponding to a spatial period $\lambda = 2.3\,a_0$.
The competition between the driving force and the landscape potential
can produce a variety of hopping patterns in the particle motion. 
The relative values of $F_{\rm ac}$, $F_{\rm dc}$, and $A_p$
control the rate and distance a  particle moves 
forward and backward in the landscape.
If $F_{\rm d}(t) > A_p$, a particle can 
overcome the barrier height of the landscape,
and 
the particle hops between minima in the energy landscape.
If the driving frequency is low,
as in Fig.~\ref{fig:2_Fd_vy_time},  
the driven particle 
moves 
in a pattern 
with the same frequency 
as the time-dependent force $F_{\rm d}(t)$,
but is modulated by the landscape period.
We explore changes in the frequency and different values of $F_{\rm ac}$ 
in Problem~\ref{ex:parameters}.
Here we vary $F_{\rm dc}$ 
while holding the remaining parameters fixed.

\begin{table}[h!]
\centering
\caption{Simulation parameters and units with comparable
  experimental values.~\cite{Juniper2015,Juniper2017} The substrate is scaled by our force units, while an experimental landscape is scaled by the Brownian motion of the particles. }%[DM changed to T xx what does $T_0 = 0$ mean?  $T_0$ is never used in the paper. xx]  }
\begin{ruledtabular}
\begin{tabular}{c c c } 
Quantity & Simulation Units & Experimental values\\
\hline
length &  $a_0 = 1$ & $ a_0 \sim 1.5 \mu m$\\
energy & $E_0 = 1$ & \\ %$ < E_0 <$ \\
%electric potential & $V(r_{ij}) = E_0/r_{ij}$ \\
%energy & $E_0 = q^2{Z^*}^2/4\pi \epsilon \epsilon_0 a_0$ \\
%dimensionless interaction strength & $q$ \\
%effective colloidal charge & $Z^*$ \\
%solvent dielectric constant & $\epsilon \epsilon_0$\\
force & $F_0 = E_0 / a_0$ & \\ %$<F_0<$\\
time &  $ \tau = \eta a_0 / F_0 $ & $ \tau \sim 3$\,s\\
velocity &  $ v_0 = a_0 / \tau $ &  $v \sim 5\,\mu$m/s \\
%driving period & $T = 100 \tau$ & $T \sim 1.3$ s \\ %frequency 0.75 Hz
%juniper uses water–ethanol mixture
substrate period & $\lambda = 2.3\,a_0$ & $\lambda = 3.5\,\mu$m \\
substrate amplitude & $A_p = 0.1\,F_0$ & $U_0 = 25 k_B T \sim 1\, {\rm J}$ \\ %$U_0 = 0.037 E_0$ 
temperature & $T = 0$  & $T \sim 290\,$K  \\
\end{tabular}
\end{ruledtabular}
\label{tab:1}
\end{table}

In Fig.~\ref{fig:2_Fd_vy_time}(a)
we plot $F_{\rm d}(t)$ 
as a function of time with the
constants $F_{\rm dc}=0.07$, 
$F_{\rm ac}=0.07$ and $f=0.01$ cycles per time unit $\tau$.
The  temporal period of the driving force is
$T = 1/f = 100\,\tau$.
In Fig.~\ref{fig:2_Fd_vy_time}(b) 
we show the $y$-position of the particle
as a function of time,
where we have
normalized $y$ by $\lambda$.
The initial particle position is $y = 0$. 
The particle moves
in the positive $y$-direction
through a distance $\Delta y = \lambda$  in the time $T$,
so that  
the average velocity 
$\langle {v}_y \rangle = \lambda f$. 
The inset of Fig.~\ref{fig:2_Fd_vy_time}(b)
shows $y$ 
over one period $100\,\tau < t < 200\,\tau$
with 
the contour plot described
in Fig.~\ref{fig:1_landscape}(a).
The motion is synchronized such that the 
driving force is a maximum when the landscape 
force is minimum,
as shown by the coincidence of the
steep slope of 
$y/\lambda$ 
and the maxima in $F_{\rm d}(t)$.
When $F_{\rm d}(t)$ is large,
the particle moves across the substrate maxima.
%shown in gray with the contour plot. said in caption

\begin{figure}[h]
\centering
\includegraphics[width=\columnwidth]{fig2_Fd_vy_time.pdf}
\caption{(a) The applied driving force $F_{\rm d}(t)$ 
  with parameters $F_{\rm dc} = 0.07$, $F_{\rm ac} = 0.07$ and $f=0.01$.
  (b) 
  The $y$-position of the driven particle
  normalized by the period of the substrate $\lambda$.
  The substrate strength is $A_p=0.1$.
  The inset   shows
  the $y$-position
  through the second period $100\,\tau<t<200\,\tau$
 along with the contour plot depicting
  the landscape potential described in Fig.~\ref{fig:1_landscape}(a).
  }
\label{fig:2_Fd_vy_time}
\end{figure}

To explore the possible hopping patterns,
we sweep through a range of $F_{\rm dc}$ for fixed $F_{\rm ac}$ and $A_p$.
In Fig.~\ref{fig:3_sweep_vyFDC} 
we increase $F_{\rm dc}$ in increments of $0.001\,F_0$,
and 
measure the average velocity $\langle v_y \rangle $ 
as a function of $F_{\rm dc}$.
We also perform the sweep for a non-oscillatory drive $F_{\rm ac} = 0$.
With no oscillating component of the driving force,
  the force-velocity relation increases monotonically 
  above the depinning threshold $F_c$ such that
  \begin{equation}
    \langle v_y \rangle \propto (F_{\rm dc}-F_c)^{-\beta},
  \end{equation}
  where the power $\beta$ varies with the
  %system type and
  type of phase transition and 
  %can
  may 
  be used to identify the universality class.\cite{Reichhardt2017} %[xx what does system type mean? xx - DM modified to phase transition]
  The critical force $F_c$ is equal to the maximum substrate force
  $A_p$.

>>>>

\begin{abstract}
  Synchronization
  plays an important role in many physical processes.
  We discuss synchronization in a 
  molecular dynamics simulation
  of a single particle
  moving through
  a viscous liquid
  while being driven 
  across a washboard potential energy landscape.
  Our results show many dynamical patterns
  as the landscape and driving force are altered.
  For certain conditions,
  the particle's velocity and location
  are synchronized or 
  phase-locked and
  form closed orbits in phase space.
  Quasi-periodic motion is common, 
  for which the
  dynamical center of motion shifts the
  phase space orbit.
  %DM removed first sentence,
  %which was strong and clear, but isolated in the abstract
  %put those ideas in this final sentence, making it a mouthful [26 words]
  By 
  isolating
  synchronized motion
  in 
  simulations and table-top experiments,
  we 
  can study %DM replaced isolate with study
  complex natural behaviors
  important to many physical processes.
 %[suggest omitting this sentence. Code not in text. We include molecular dynamics code
 % to simulate and characterize the
  %particle dynamics.] 
\end{abstract}

@@

\section{Introduction} 
Synchronization is a universal phenomena
in which
individual oscillators change frequency due
to external stimuli.\cite{Pikovsky2003}
The
flickering patterns of
candle flames 
mediated by temperature fluctuations,\cite{Okamoto2016}
the vibrations of singing wine glasses interacting 
through sound waves,\cite{Arane2009}
and metronomes vibrating through a supporting platform\cite{Jia2015}
are examples of in-phase environmentally coupled oscillations. 
Biological systems benefit from cooperative
synchronization --
birds  coordinate their wing flaps
to optimize energy use during flight,\cite{Portugal2014}
frogs alternate croaking patterns,\cite{Aihara2014}
humans clap in time with music,\cite{Tranchant2016}
and 
at a cellular level, 
neurons simultaneously fire in cardiac muscle\cite{MartinHall1999}
and brain tissue.\cite{Singer1999}
External forcing
can cause or regulate synchronization.
%where external forces
%add or remove energy from the system.
For example, 
an electrical pacemaker 
regulates a heart beat and
a pulsed light modifies the
flashing pattern of fireflies.

The model we discuss
%DM add
in this paper 
%DM end
resembles an externally forced pendulum.
%damping
%transition
%%DM - focus more on single particle with environment
A single pendulum 
can swing in time with a %be driven or damped by
driving force
like  %a skateboarder in a half-pipe or
a pushed child on a swing,
and may be damped
by mechanical friction.
%of gravity to return
%to an equilibrium position,
%or  to decreasing amplitude
% or air resistance.
We focus on 
synchronized phase-locking or mode-locking,
which 
first appeared in the scientific literature
in 1665 with 
Huygens'  experiments on the
motions of %synchronized 
pendula in wall-mounted clocks.\cite{Bennett2002}
A locked-mode is an integer frequency ratio
in coupled oscillators. %Clock pendula swing in phase
%when coupled through shared vibrations.
In Huygens' clocks,
the mode-locked pendula were observed to swing at the same rate
in the same direction (1:1 mode) or in
opposite directions ($-1$:1 mode).
Higher modes might be achieved
by altering the pendula lengths --
a 2:1 mode occurs when a simple pendulum
swings with twice the period
of another one quarter the length.
%DM transition sentence
%Many subsequent 
Simulations and laboratory experiments
have revealed 
mode-locked synchronization
in a diverse range of 
oscillation phenomena.
%Single oscillators 
%can demonstrate a variety of synchronized behaviors.
%Colloid particles in optical traps are an excellent
%model system to explore %these 
%the dynamics of mode-locking. %synchronization.

Colloids 
provide an ideal system to 
control and study
phase-locked motions
%are useful for
%understanding synchronization
at a single particle level.\cite{Juniper2015,Juniper2017}
% such as synchronized mode-locking can be studied with
%An advantage of colloid studies
In experiments %DM moved the ``in experiments'' to second sentence
typical colloids are   
plastic spheres suspended in
de-ionized water or silica beads suspended in organic solvent.
Colloids can be trapped
with 
radiation pressure from 
a laser beam.\cite{Ashkin1997} 
A colloid centered in an optical trap is 
uniformly bombarded by photons
and oscillates
about a equilibrium point.
Off-center colloids 
experience a net force
due to uneven photon collisions across
the particle's surface.
Diffraction gratings can be used to create
more complex light environments, 
such as periodic patterns of minima
suitable for synchronization studies.\cite{Grier2003}  
A driven colloid
can demonstrate
synchronized
motion within one trap,
or synchronized hopping between multiple traps.
%DM moved this sentence up for better flow
In the following simulations, we confine colloids  
to move along a periodic pattern of 
minima and maxima defined by a sine wave.

Although many systems oscillate,
trapped colloids are useful for 
providing controlled 
measurements of the
step-by-step dynamics of the particle motion.
Because
colloids are large 
and move slowly,  the positions of the
particles
can be measured in real time 
with an optical camera.\cite{Pertsinidis2001}
Simulations provide complementary
information,
where environmental and driving parameters may be precisely
and quickly modified to explore
the many oscillation patterns.

A colloid driven across a periodic
substrate may 
synchronize its location
to the pattern of the external drive.
%moving back and forth in time with
%the beat, or
%moving between substrate minima.
When 
a constant or 
dc drive is applied,
the particle's velocity is modulated by 
the potential energy landscape exerted on the particle.   
Below some threshold  
the dc force is not strong enough to push the particle
across a potential maximum so the average velocity is zero,
a phenomena referred to as pinning.\cite{Reichhardt2017}
Above the pinning threshold,
a dc driven particle
%a particle subject to a constant
%drive force
increases its speed at a rate proportional
to the external drive.  
When the applied force varies periodically,
an ac drive,
%cause
the particle can hop back and forth across
the landscape minima.
When a particle is driven across a substrate
by combined 
ac and dc forces,
it is possible for the particle to demonstrate 
mode-locking.
In these modes,  
%Due to the combination of external drive
%and substrate minima,
the particle hops at a controlled rate such that 
the average particle velocity is 
%of an  driven particle 
fixed for a range of dc drive forces
with a magnitude proportional to the
landscape period and driving frequency.\cite{Reichhardt2015}
%Many synchronized patterns occur,
%controlled by the substrate pattern and driving force.

In this paper we discuss simulations 
of the mode-locked dynamics
of a confined particle driven over
a washboard shaped potential energy landscape.
The model is presented in Sec.~\ref{sec:MD} and 
our results are summarized in
%including the synchronized motion of a single confined particle
%driven across a periodic landscape in 
Sec.~\ref{sec:results}. In Sec.~\ref{sec:conclusion}
we relate  our results 
to a broad variety of physical systems beyond 
optically trapped colloids.
%and superconducting vortices in the presence of periodic pinning arrays.
We include problems for interested readers 
in Sec.~\ref{sec:problems}.

\section{Simulation}
\label{sec:MD}
We use a classical model for 
studying the dynamics, 
using the net force on a particle to calculate
its trajectory.
The particle is confined in a one-dimensional 
box of length $L=46.\bar{6}\,a_0$,
where $a_0$ is a dimensionless unit of length.
(In the simulations we set $a_0 = 1$.)
The particle 
has
position $\vec{r} = y \hat{y}$ 
and velocity $\vec{v} = d\vec{r}/dt$.
The edges of the system are treated by
periodic boundary conditions,
such that a particle leaving the edge of the system is mapped
back to a position within the simulation boundaries 
by the transformation $y+L \rightarrow y$. 
We show a schematic of the system in Fig.~\ref{fig:1_landscape}(a).
The units of the simulated variables are summarized in Table~\ref{tab:1}.

We confine the particles using a position dependent 
potential energy, called a landscape or substrate.
The landscape is modulated in the $y$-direction
by the periodic function 
 \begin{equation}
   U(y) = U_0 \cos{(2 \pi y / \lambda)},
     \label{eq:ysubstrate}
\end{equation}
where $\lambda=L/N_p$, with $N_p$ equal to the number of periods,
and $U_0$ is a parameter
 that sets the depth of the minima
 with  units of energy $E_0$. 
 We plot this function in 
 Fig.~\ref{fig:1_landscape}
 for $N_p = 3$.  In Fig.~\ref{fig:1_landscape}(a) we show 
 the $x$-$y$ plane with a contour plot of $U(y)$ 
 to illustrate
 the two-dimensional potential energy landscape;
 the maxima are colored gray and the minima are colored white.  
 Because the motion is one-dimensional, the particle only moves along the $y$-direction in the two-dimensional plane. %[xx added this sentence because the reader might be confused that you stated that the particle is confined to one dimension xx]

The confining force on the particle 
is given by
 \begin{equation}
 \vec{F}_{\ell}(\vec{r}) = - \vec \nabla U(\vec{r}),
 \label{eq:dudr}
 \end{equation}
where $\ell$ denotes the landscape.
 In Fig.~\ref{fig:1_landscape}(b) we plot  
 $U(y)$ to illustrate how the magnitude
 $|\vec{F}_{\ell}|$ is calculated from the particle position $y$.

@@

The equation of motion  yields the velocity
of a particle at position $\vec{r}$ 
and simulation time $t$. At each time step
we evaluate the net force on the particle as a function of its position
$\vec{r}(t)$
and then integrate
the equation of motion to move the particle
to its updated position.
Because the acceleration is zero,
the integration of the equation of motion
is performed using 
the simple first-order Euler method 
\begin{equation}
  \vec{r}(t+\Delta t) = \vec{r}(t) + \vec{v}(t) \Delta t
    \label{eq:euler}
\end{equation}
for a time step $\Delta t = 0.1\,\tau$.
In 
Problem~\ref{ex:euler}
we describe a program for 
treating differential equations such as Eq.~\eqref{eq:euler}. %[xx note change xx]

\begin{figure} % [h!]
\centering
\includegraphics[width=\columnwidth]{fig1_landscape.pdf}
\caption{
Schematic of the simulation of a single particle
  driven across a washboard potential energy landscape.
  (a) View of the $x$-$y$ plane. 
  The time-dependent applied driving force $\vec{F}_{\rm d}$
  is parallel to the $y$-axis.
  The landscape is shown with
  maxima of the potential energy in gray 
  and minima in white. 
  The particle is 
  subject to the competing forces of the landscape and the applied driving force.
  (b) The potential energy function $U(y)$. 
  The particle in (a) is shown at the same $y$-position in (b).
  The slope of $U(y)$ is the 
  magnitude of the force $\vec{F}_{\ell}$. 
  }
\label{fig:1_landscape}
\end{figure}

@@

\label{sec:results}
%Here we drive 
%a single particle 
%across the landscape. 
The numerical implementation of the force due to the landscape 
is calculated using Eqs.~\eqref{eq:ysubstrate} and \eqref{eq:dudr} with
\begin{equation}
  \label{eq:force}
  {F_{\ell}}_y(y) = A_{p} \sin{(2 \pi y / \lambda)},
\end{equation}
where the force is scaled by the parameter $A_{p} = 2\pi U_0/\lambda$.
In this section we fix the landscape parameters
to $A_{p} = 0.1\,F_0$ 
with $N_p=20$, 
corresponding to a spatial period $\lambda = 2.3\,a_0$.
The competition between the driving force and the landscape potential
can produce a variety of hopping patterns in the particle motion. 
The relative values of $F_{\rm ac}$, $F_{\rm dc}$, and $A_p$
control the rate and distance a  particle moves 
forward and backward in the landscape.
If $F_{\rm d}(t) > A_p$, a particle can 
overcome the barrier height of the landscape,
and 
the particle hops between minima in the energy landscape.
If the driving frequency is low,
as in Fig.~\ref{fig:2_Fd_vy_time},  
the driven particle 
moves 
in a pattern 
with the same frequency 
as the time-dependent force $F_{\rm d}(t)$,
but is modulated by the landscape period.
We explore changes in the frequency and different values of $F_{\rm ac}$ 
in Problem~\ref{ex:parameters}.
Here we vary $F_{\rm dc}$ 
while holding the remaining parameters fixed.

\begin{table}[h!]
\centering
\caption{Simulation parameters and units with comparable
  experimental values.~\cite{Juniper2015,Juniper2017} The substrate is scaled by our force units, while an experimental landscape is scaled by the Brownian motion of the particles. }
\begin{ruledtabular}
\begin{tabular}{c c c } 
Quantity & Simulation units & Experimental values\\
\hline
length &  $a_0 = 1$ & $ a_0 \sim 1.5 \mu m$\\
energy & $E_0 = 1$ & \\ %$ < E_0 <$ \\
%electric potential & $V(r_{ij}) = E_0/r_{ij}$ \\
%energy & $E_0 = q^2{Z^*}^2/4\pi \epsilon \epsilon_0 a_0$ \\
%dimensionless interaction strength & $q$ \\
%effective colloidal charge & $Z^*$ \\
%solvent dielectric constant & $\epsilon \epsilon_0$\\
force & $F_0 = E_0 / a_0$ & \\ %$<F_0<$\\
time &  $ \tau = \eta a_0 / F_0 $ & $ \tau \sim 3$\,s\\
velocity &  $ v_0 = a_0 / \tau $ &  $v \sim 5\,\mu$m/s \\
%driving period & $T = 100 \tau$ & $T \sim 1.3$ s \\ %frequency 0.75 Hz
%juniper uses water–ethanol mixture
substrate period & $\lambda = 2.3\,a_0$ & $\lambda = 3.5\,\mu$m \\
substrate amplitude & $A_p = 0.1\,F_0$ & $U_0 = 25 k_B T \sim 1\, {\rm J}$ \\ %$U_0 = 0.037 E_0$ 
temperature & $T = 0$  & $T \sim 290\,$K  \\
\end{tabular}
\end{ruledtabular}
\label{tab:1}
\end{table}

In Fig.~\ref{fig:2_Fd_vy_time}(a)
we plot $F_{\rm d}(t)$ 
as a function of time with $F_{\rm dc}=0.07$, 
$F_{\rm ac}=0.07$, and $f=0.01$ cycles per time unit $\tau$.
The  temporal period of the driving force is
$T = 1/f = 100\,\tau$.
In Fig.~\ref{fig:2_Fd_vy_time}(b) 
we show the $y$-position of the particle
as a function of time,
where we have
normalized $y$ by $\lambda$.
The initial particle position is $y = 0$. 
The particle moves
in the positive $y$-direction
through a distance $\Delta y = \lambda$  in the time $T$,
so that  
the average velocity 
$\langle {v}_y \rangle = \lambda f$. 
The inset of Fig.~\ref{fig:2_Fd_vy_time}(b)
shows $y$ 
over one period $100\,\tau < t < 200\,\tau$
with 
the contour plot described
in Fig.~\ref{fig:1_landscape}(a).
The motion is synchronized such that the 
driving force is a maximum when the landscape 
force is minimum,
as shown by the coincidence of the
steep slope of 
$y/\lambda$ 
and the maxima in $F_{\rm d}(t)$.
When $F_{\rm d}(t)$ is large,
the particle moves across the substrate maxima.

\begin{figure}[h]
\centering
\includegraphics[width=\columnwidth]{fig2_Fd_vy_time.pdf}
\caption{(a) The applied driving force $F_{\rm d}(t)$ 
  with  $F_{\rm dc} = 0.07$, $F_{\rm ac} = 0.07$, and $f=0.01$.
  (b) 
  The $y$-position of the  particle
  normalized by the period of the substrate $\lambda$.
  The substrate strength is $A_p=0.1$.
  The inset   shows
  the $y$-position
  through the second period $100\,\tau<t<200\,\tau$
 along with the contour plot depicting
  the landscape potential shown in Fig.~\ref{fig:1_landscape}(a).
  }
\label{fig:2_Fd_vy_time}
\end{figure}

To explore the possible hopping patterns,
we sweep through a range of values of $F_{\rm dc}$ for fixed $F_{\rm ac}$ and $A_p$.
In Fig.~\ref{fig:3_sweep_vyFDC} 
we increase $F_{\rm dc}$ in increments of $0.001\,F_0$,
and 
measure the average velocity $\langle v_y \rangle $ 
as a function of $F_{\rm dc}$.
We also perform the sweep for a non-oscillatory drive $F_{\rm ac} = 0$.
With no oscillating component of the driving force,
  the force-velocity relation increases monotonically 
  above the depinning threshold $F_c$ such that
  \begin{equation}
    \langle v_y \rangle \propto (F_{\rm dc}-F_c)^{-\beta},
  \end{equation}
  where the power $\beta$ varies with the
  type of phase transition and 
  %can
  may 
  be used to identify the universality class.\cite{Reichhardt2017} 
  The critical force $F_c$ is equal to the maximum substrate force
  $A_p$.
  % [xx would be good to follow up on your mention of universality class in a problem. The exponent beta is usually associated with the decrease of the order parameter, not a divergence xx]
  %DM - in my experience, we use beta for this parameter.  For example
  %Dynamic regimes for driven colloidal particles on a periodic substrate at commensurate and incommensurate fillings D McDermott, J Amelang, CJO Reichhardt, C Reichhardt, Physical Review E 88 (6), 062301

The addition of an ac drive leads
  to the formation of modes.
  A mode is a periodic pattern of hops
  with a constant average particle velocity $\langle {v}_{y} \rangle$
  over a range of driving forces $F_{\rm dc}$.
  In Fig.~\ref{fig:3_sweep_vyFDC}
  we sweep $F_{\rm dc}$
  with $F_{\rm ac} = 0.07$ and $f=0.01$.
Each value of $F_{\rm dc}$ yields a different pattern of hops
between substrate minima
performed by the particle
due to the landscape confinement.  %xx note change xx]
At 
low values of $F_{\rm dc}$, the average velocity $\langle v_y \rangle$ is zero.
Because 
$A_p$ is large compared to the extrema of $F_{\rm d}(t)$,
the particle oscillates back and forth
in a single minima with no net velocity, an example of 
a 0:0 mode.
At higher values of $F_{\rm dc}$,    
$\langle v_{y} \rangle$ increases in steps of uniform height,
$\langle v_{y} \rangle = n \lambda f$,
where $n$ is an integer.
We observe a mode of $n=1$
for   $0.05 < F_{\rm dc} < 0.08$,
$n=2$ for $0.08 < F_{\rm dc} < 0.11$,
$n=3$ for $0.12 < F_{\rm dc} < 0.13$,
$n=4$ for $0.14 < F_{\rm dc} < 0.155$, and 
$n=5$ for $0.155 < F_{\rm dc} < 0.16$.
Higher modes are not visible.
The step width is nonlinear
and depends on the strength of $F_{\rm ac}$
for this landscape potential.\cite{Reichhardt2000,Juniper2017}
These steps, known as Shapiro steps, can exhibit a variety of
interesting patterns
such as a devil's staircase related to chaotic dynamics.\cite{Bak1986}
